---
title: "analysis plan"
author: "Jack Yan"
date: "5/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # data manipulation
library(corrplot) # correlation plot
library(caret) # model training

# library(AppliedPredictiveModeling)
```

```{r, warning=F}
dat = 
  read_csv("./data.csv") %>% 
  select(-id, -X33) 
dim(dat)
```

# Create a test dataset
```{r, eval=FALSE}
set.seed(123123)
test_dat = sample_n(dat, 569/5)
train_dat = anti_join(dat, test_dat)
write_csv(train_dat, "train.csv")
write_csv(test_dat, "test.csv")
```

# Load train and test data
```{r}
train_df = 
  read_csv("./train.csv") %>% 
  mutate(diagnosis = as_factor(diagnosis))
test_df = 
  read_csv("./test.csv")  %>% 
  mutate(diagnosis = as_factor(diagnosis))
```

# Exploratory data analysis

### Clustering (Unsupervised learning)
```{r}

```

### Correlation plots
```{r}
x <- model.matrix(diagnosis~., train_df)[,-1]
y <- train_df$diagnosis
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5) 
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1) 
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2) 
trellis.par.set(theme1)
# data_rename = 
#   data %>% 
#   rename(income = income_composition_of_resources,
#          per_expend = percentage_expenditure,
#          to_expend = total_expenditure,
#          death5 = under_five_deaths,
#          thin5_9 = thinness_5_9_years,
#          thin1_19 = thinness_1_19_years)
x_rename <- model.matrix(life_expectancy~., data_rename)[,-1]

par(cex = 0.7)
corrplot(cor(x), tl.srt = 45, order = 'hclust', type = 'upper')
```

### One-to-one relation between classes and covariates
```{r}
# Distribution of response classes with regard to each variable
transparentTheme(trans = .4)
featurePlot(x = train_df[,2:11], 
            y = train_df$diagnosis,
            scales = list(x = list(relation="free"), 
                          y = list(relation="free")),
            plot = "density", pch = "|",
            auto.key = list(columns = 2))
```

# Model building, assessing performance, and variable importance

## Linear methods
### Logistic regression
### regularized logistic regression (glmnet)
### LDA

## Non-linear methods
### QDA
### Naive Bayes
### KNN

## Classification Trees and Ensemble methods
### a single tree
### bagging
### random forests
### boosting (Adaboosting and binomial loss function)

# Variable importance, Visualization, and interpretation (L7_2.rmd): PDP, ICE, etc.

# Model comparison

# Final model interpretation


